---
title: "01models"
output: html_document
---

```{r setup, include=FALSE}
library(miceadds)
library(lmtest)
library(multiwayvcov)
library(modelsummary)
library(fabricatr)
library(mediation)
library(ivpack)
library(tidyr)
library(lubridate)
library(dplyr)
library(stargazer)
library(plm)
library(lfe)
library(ggplot2)
library(forcats)
library(scales)
```

# 1. VMT model
# 1.1 load data
```{r}
data_vmt0<-read.csv('data_processed/data_vmt.csv')
data_vmt0$date<-as.Date(data_vmt0$date)
data_vmt0$month<-month(data_vmt0$date)
data_vmt0$year<-year(data_vmt0$date)
data_vmt<-data_vmt0%>%
  filter((date>as.Date('2020-03-01'))&(date<as.Date('2022-04-01')))
data_vmt
```
# 1.2 fixed-effects regression
```{r}
ols_vmt<-lm(vmt_19pct~onsite_pct+reopened+net_mig_rate+covid_cases+vaccine+vrm_19pct+log(pop_million)+factor(state)+factor(month),data=data_vmt)
rob.ols_vmt<-coeftest(ols_vmt,function(x){cluster.vcov(x, data_vmt$state)})
```

# 1.3 first-stage 2SLS regression
```{r}
st_state1<-lm(onsite_pct~iv+reopened+net_mig_rate+covid_cases+vaccine+vrm_19pct+log(pop_million)+factor(state)+factor(month),data=data_vmt)
rob.st_state1<-coeftest(st_state1,vcov = vcovHC(st_state1,type = "HC3"))

summary(st_state1)$r.squared
st_state1_pred<-st_state1$fitted.values
data_vmt$st1_pred<-st_state1_pred
```
# 1.4 second-stage 2SLS regression
```{r}
# 1.4.1 main
st2_st_main<-lm(vmt_19pct~st1_pred+reopened+net_mig_rate+covid_cases+vaccine+vrm_19pct+log(pop_million)+factor(month)+factor(state),data=data_vmt)

rob.st2_st_main<-coeftest(st2_st_main,vcov = vcovHC(st2_st_main,type = "HC3"))

# 1.4.2 spatial heteogeneity
st2_st_trans<-lm(vmt_19pct~st1_pred+I(st1_pred*log(transitcommute_pct))+reopened+net_mig_rate+covid_cases+vaccine+vrm_19pct+log(pop_million)+factor(month)+factor(state),data=data_vmt)

rob.st2_st_trans<-coeftest(st2_st_trans,vcov = vcovHC(st2_st_trans,type = "HC3"))

# 1.4.3 time heteogeneity
st2_st_time<-lm(vmt_19pct~I(st1_pred*((date>=as.Date('2020-04-01'))&(date<as.Date('2020-07-01'))))+I(st1_pred*((date>=as.Date('2020-07-01'))&(date<as.Date('2020-10-01'))))+I(st1_pred*((date>=as.Date('2020-10-01'))&(date<as.Date('2021-01-01'))))+I(st1_pred*((date>=as.Date('2021-01-01'))&(date<as.Date('2021-04-01'))))+I(st1_pred*((date>=as.Date('2021-04-01'))&(date<as.Date('2021-07-01'))))+I(st1_pred*((date>=as.Date('2021-07-01'))&(date<as.Date('2021-10-01'))))+I(st1_pred*((date>=as.Date('2021-10-01'))&(date<as.Date('2022-01-01'))))+I(st1_pred*((date>=as.Date('2022-01-01'))&(date<as.Date('2022-04-01'))))+reopened+net_mig_rate+covid_cases+vaccine+vrm_19pct+log(pop_million)+factor(month)+factor(state),data=data_vmt)

rob.st2_st_time<-coeftest(st2_st_time,vcov = vcovHC(st2_st_time,type = "HC3"))

```

# 1.5 calculate CO2 emissions
# 1.5.1 load VMT-CO2 data
```{r}
co2_df<-read.csv('auxiliary_data/co2_lbs_mile2017.csv')
co2_df<-co2_df%>%
  select(state,co2_lbs_mile)
```
# 1.5.2 calculate CO2 effect for each state
```{r}
results_st <- tidy(rob.st2_st_trans)
results_st

cov_inter<-cluster.vcov(st2_st_trans, data_vmt$state)['st1_pred','I(st1_pred * log(transitcommute_pct))']
var_main<-cluster.vcov(st2_st_trans, data_vmt$state)['st1_pred','st1_pred']
var_2<-cluster.vcov(st2_st_trans, data_vmt$state)['I(st1_pred * log(transitcommute_pct))','I(st1_pred * log(transitcommute_pct))']

vmt19_avg<-data_vmt%>%
  filter(year==2021)%>%
  group_by(state)%>%
  dplyr::summarise(vmt19=mean(vmt19))

```
```{r}
df_effect<-data_vmt%>%
  select(state,transitcommute_pct,iv)%>%
  group_by(state)%>%
  dplyr::summarise(transitcommute_pct=mean(transitcommute_pct),iv=mean(iv))%>%
  left_join(vmt19_avg,by='state')%>%
  mutate(l_trans=log(transitcommute_pct))%>%
  mutate(Coefficient=rob.st2_st_trans['st1_pred','Estimate']+rob.st2_st_trans['I(st1_pred * log(transitcommute_pct))','Estimate']*log(transitcommute_pct))%>%
  mutate(SE=sqrt(var_main+var_2*(log(transitcommute_pct)**2)+2*cov_inter*log(transitcommute_pct)))%>%
  mutate(Coef.low_95=Coefficient-1.96*SE,Coef.high_95=Coefficient+1.96*SE)%>%
  mutate(vmt_effect=Coefficient*vmt19/100,vmt.low_95=Coef.low_95*vmt19/100,vmt.high_95=Coef.high_95*vmt19/100)%>%
  left_join(co2_df,by='state')%>%
  mutate(co2_effect=vmt_effect*co2_lbs_mile,co2.low_95=vmt.low_95*co2_lbs_mile,co2.high_95=vmt.high_95*co2_lbs_mile)%>%
  arrange(desc(co2_effect))

df_effect<-df_effect%>%
  mutate(co2_effect_yr=co2_effect*12)%>%
  mutate(co2.low_95_yr=co2.low_95*12,co2.high_95_yr=co2.high_95*12)
```
# 1.5.3 plot CO2 effect for each state
```{r}

df_effect %>%
  mutate(state = fct_reorder(state, co2_effect))%>%
ggplot(aes(x = state, y = co2_effect)) +
        geom_hline(yintercept = 0,
                   colour = gray(1/2), lty = 2) +
        geom_bar(stat="identity",fill='steelblue4') + 
        geom_linerange(aes(x = state, 
                     ymin = co2.low_95,
                     ymax = co2.high_95),
                   lwd = 0.7,position = position_dodge(width = 0),color='orange2')+
  geom_text(aes(label = round(co2_effect,1)),hjust = -0.5,size = 7) + 
        coord_flip()+
    labs(y=bquote('Effect on monthly' ~CO[2]*' emissions (million pounds)'))+
  labs(x="State")+
  theme(legend.key=element_blank(),
        axis.ticks=element_blank(),
        axis.text.x=element_text(size=20,family="Arial"), 
        axis.text.y=element_text(size=20,family="Arial"), 
        axis.title=element_text( size=25, family="Arial"),
        axis.title.y = element_text(margin = margin(r = 12)),
        axis.title.x = element_text(margin = margin(t = 10)),
        legend.text=element_text(size=20, family="Arial",margin = margin(t = 7)), 
        panel.background=element_blank(),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.grid.major.x = element_line(colour = "lightgrey"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_line(colour = "lightgrey"),
        panel.grid.minor.y = element_blank(),
        # panel.grid.minor = element_line(colour = "lightgrey"),
        legend.box.background = element_rect(colour = "black"),
        text = element_text(size = 20),
        plot.margin = margin(t = 0.5,  # Top margin
                             r = 0.5,  # Right margin
                             b = 0.2,  # Bottom margin
                             l = 0.2, "cm"))

df_effect %>%
  mutate(state = fct_reorder(state, co2_effect_yr))%>%
ggplot(aes(x = state, y = co2_effect_yr)) +
        geom_hline(yintercept = 0,
                   colour = gray(1/2), lty = 2) +
        geom_bar(stat="identity",fill='steelblue4') + 
        geom_linerange(aes(x = state, 
                     ymin = co2.low_95_yr,
                     ymax = co2.high_95_yr),
                   lwd = 0.7,position = position_dodge(width = 0),color='orange2')+
  geom_text(aes(label = round(co2_effect_yr,1)),hjust = -0.5,size = 7) + 
        coord_flip()+
    labs(y=bquote('Effect on annual' ~CO[2]*' emissions (million pounds)'))+
  labs(x="State")+
  theme(legend.key=element_blank(),
        axis.ticks=element_blank(),
        axis.text.x=element_text(size=20,family="Arial"), 
        axis.text.y=element_text(size=20,family="Arial"), 
        axis.title=element_text( size=25, family="Arial"),
        axis.title.y = element_text(margin = margin(r = 12)),
        axis.title.x = element_text(margin = margin(t = 10)),
        legend.text=element_text(size=20, family="Arial",margin = margin(t = 7)), 
        panel.background=element_blank(),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.grid.major.x = element_line(colour = "lightgrey"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_line(colour = "lightgrey"),
        panel.grid.minor.y = element_blank(),
        # panel.grid.minor = element_line(colour = "lightgrey"),
        legend.box.background = element_rect(colour = "black"),
        text = element_text(size = 20),
        plot.margin = margin(t = 0.5,  # Top margin
                             r = 0.5,  # Right margin
                             b = 0.2,  # Bottom margin
                             l = 0.2, "cm"))
# ggsave('plots/co2_yr.png', dpi = 800, width = 12, height = 16, units = "in")
```
# 1.5.4 calculate the aggregate CO2 effect
```{r}
total_us_co2=1939.46-11.9-12.3

##VMT-related CO2 (yearly)
vmt_co2_19=sum(df_effect$vmt19*df_effect$co2_lbs_mile)*12
# sum(df_effect$co2_effect)*12/vmt_co2_19

##use official transport-sector emission as the baseline
co2_sum=sum(df_effect$co2_effect)*12/2204.62 #MtCO2

print(paste0('1% increase in remote work will cause reduction in annual CO2 emission: ',round(co2_sum,2),' MtCO2, ',round(co2_sum*100/total_us_co2,2),'% of the 2019 level'))
```

# 1.6 plot the coefficient for each state
```{r}
df_effect %>%
  mutate(state = fct_reorder(state, Coefficient))%>%
ggplot(aes(x = state, y = Coefficient)) +
        geom_hline(yintercept = 0,
                   colour = gray(1/2), lty = 2) +
        geom_point(aes(x = state,
                    y = Coefficient),position = position_dodge(width = 0),size=3) +
        geom_linerange(aes(x = state,
                     ymin = Coef.low_95,
                     ymax = Coef.high_95),
                   lwd = 1,position = position_dodge(width = 0)) +
        coord_flip()+
  labs(x="State")+
  theme(legend.key=element_blank(),
        axis.ticks=element_blank(),
        axis.text.x=element_text(size=20,family="Arial"), 
        axis.text.y=element_text(size=20,family="Arial"), 
        axis.title=element_text( size=25, family="Arial"),
        axis.title.y = element_text(margin = margin(r = 12)),
        axis.title.x = element_text(margin = margin(t = 10)),
        legend.text=element_text(size=20, family="Arial",margin = margin(t = 7)), 
        panel.background=element_blank(),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.grid.major.x = element_line(colour = "lightgrey"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_line(colour = "lightgrey"),
        panel.grid.minor.y = element_blank(),
        # panel.grid.minor = element_line(colour = "lightgrey"),
        legend.box.background = element_rect(colour = "black"),
        text = element_text(size = 20),
        plot.margin = margin(t = 0.5,  # Top margin
                             r = 0.5,  # Right margin
                             b = 0.2,  # Bottom margin
                             l = 0.2, "cm"))
ggsave('plots/coef_vmt.png', dpi = 800, width = 12, height = 16, units = "in")
```

# 2. Congestion model
# 2.1 load data
```{r}
data_cong0<-read.csv('data_processed/data_congestion.csv')
data_cong0$date<-as.Date(data_cong0$date)
data_cong0$month<-month(data_cong0$date)
data_cong0$year<-year(data_cong0$date)
data_cong<-data_cong0%>%
  filter((date>as.Date('2020-03-01'))&(date<as.Date('2022-01-01')))
data_cong
```
# 2.2 fixed-effects regression
```{r}
ols_cong<-lm(congestion_19pct~onsite_pct+S+reopened+net_mig_rate+covid_cases+vaccine+vrm_19pct+log(pop_million)+factor(month)+factor(MSA.Code),data=data_cong)
rob.ols_cong<-coeftest(ols_cong,vcov = vcovHC(ols_cong,type = "HC3"))

```
# 2.3 first-stage 2SLS regression
```{r}
st_cong1<-lm(onsite_pct~iv+S+reopened+net_mig_rate+covid_cases+vaccine+vrm_19pct+log(pop_million)+factor(month)+factor(MSA.Code),data=data_cong)

summary(st_cong1)$r.squared
st_cong1_pred<-st_cong1$fitted.values
data_cong$st1_pred<-st_cong1_pred

rob.st_cong1<-coeftest(st_cong1,vcov = vcovHC(st_cong1,type = "HC3"))
```
# 2.4 second-stage 2SLS regression
```{r}
# 2.4.1 main
st2_cong<-lm(congestion_19pct~st1_pred+S+reopened+net_mig_rate+covid_cases+vaccine+vrm_19pct+log(pop_million)+factor(month)+factor(MSA.Code),data=data_cong)

rob.st2_cong<-coeftest(st2_cong,vcov = vcovHC(st2_cong,type = "HC3"))

# 2.4.2 spatial heterogeneity
st2_cong_pop<-lm(congestion_19pct~st1_pred+I(st1_pred*log(pop_million))+S+reopened+net_mig_rate+covid_cases+vaccine+vrm_19pct+log(pop_million)+factor(month)+factor(MSA.Code),data=data_cong)
rob.st2_cong_pop<-coeftest(st2_cong_pop,vcov = vcovHC(st2_cong_pop,type = "HC3"))

# 2.4.3 time heterogeneity
st2_cong_time<-lm(congestion_19pct~I(st1_pred*((date>=as.Date('2020-04-01'))&(date<as.Date('2020-07-01'))))+I(st1_pred*((date>=as.Date('2020-07-01'))&(date<as.Date('2020-10-01'))))+I(st1_pred*((date>=as.Date('2020-10-01'))&(date<as.Date('2021-01-01'))))+I(st1_pred*((date>=as.Date('2021-01-01'))&(date<as.Date('2021-04-01'))))+I(st1_pred*((date>=as.Date('2021-04-01'))&(date<as.Date('2021-07-01'))))+I(st1_pred*((date>=as.Date('2021-07-01'))&(date<as.Date('2021-10-01'))))+I(st1_pred*((date>=as.Date('2021-10-01'))&(date<as.Date('2022-01-01'))))+S+reopened+net_mig_rate+covid_cases+vaccine+vrm_19pct+log(pop_million)+factor(month)+factor(MSA.Code),data=data_cong)

rob.st2_cong_time<-coeftest(st2_cong_time,vcov = vcovHC(st2_cong_time,type = "HC3"))
```

# 2.5 plot the coefficient for each MSA
```{r}
results_cong <- tidy(rob.st2_cong_pop)
results_cong

cov_inter<-cluster.vcov(st2_cong_pop, data_cong$MSA.Code)['st1_pred','I(st1_pred * log(pop_million))']
var_main<-cluster.vcov(st2_cong_pop, data_cong$MSA.Code)['st1_pred','st1_pred']
var_2<-cluster.vcov(st2_cong_pop, data_cong$MSA.Code)['I(st1_pred * log(pop_million))','I(st1_pred * log(pop_million))']

df_effect_cong<-data_cong%>%
  filter(date==as.Date('2020-04-01'))%>%
  select(MSA.Name,MSA.Code,pop_million)%>%
  mutate(Coefficient=rob.st2_cong_pop['st1_pred','Estimate']+rob.st2_cong_pop['I(st1_pred * log(pop_million))','Estimate']*log(pop_million))%>%
  mutate(SE=sqrt(var_main+var_2*(log(pop_million)**2)+2*cov_inter*log(pop_million)))%>%
  mutate(Coef.low_95=Coefficient-1.96*SE,Coef.high_95=Coefficient+1.96*SE)
```

```{r}
df_effect_cong %>%
  mutate(MSA.Name = fct_reorder(MSA.Name, Coefficient))%>%
ggplot(aes(x = MSA.Name, y = Coefficient)) +
        geom_hline(yintercept = 0,
                   colour = gray(1/2), lty = 2) +
        geom_point(aes(x = MSA.Name,
                    y = Coefficient),position = position_dodge(width = 0),size=3) +
        geom_linerange(aes(x = MSA.Name,
                     ymin = Coef.low_95,
                     ymax = Coef.high_95),
                   lwd = 1,position = position_dodge(width = 0)) +
        coord_flip()+
  labs(x="MSA")+
  theme(legend.key=element_blank(),
        axis.ticks=element_blank(),
        axis.text.x=element_text(size=20,family="Arial"), 
        axis.text.y=element_text(size=20,family="Arial"), 
        axis.title=element_text( size=25, family="Arial"),
        axis.title.y = element_text(margin = margin(r = 12)),
        axis.title.x = element_text(margin = margin(t = 10)),
        legend.text=element_text(size=20, family="Arial",margin = margin(t = 7)), 
        panel.background=element_blank(),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.grid.major.x = element_line(colour = "lightgrey"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_line(colour = "lightgrey"),
        panel.grid.minor.y = element_blank(),
        # panel.grid.minor = element_line(colour = "lightgrey"),
        legend.box.background = element_rect(colour = "black"),
        text = element_text(size = 20),
        plot.margin = margin(t = 0.5,  # Top margin
                             r = 0.5,  # Right margin
                             b = 0.2,  # Bottom margin
                             l = 0.2, "cm"))
ggsave('plots/coef_cong.png', dpi = 800, width = 12, height = 16, units = "in")
```
# 3. Transit model
# 3.1 load data
```{r}
data_trans0<-read.csv('data_processed/data_transit.csv')
data_trans0$date<-as.Date(data_trans0$date)
data_trans0$month<-month(data_trans0$date)
data_trans0$year<-year(data_trans0$date)
data_trans<-data_trans0%>%
  filter((date>as.Date('2020-03-01'))&(date<as.Date('2022-04-01')))
data_trans
```
# 3.2 fixed-effects regression
```{r}
ols_trans<-lm(ridership_19pct~onsite_pct+reopened+net_mig_rate+covid_cases+vaccine+vrm_19pct+log(pop_million)+factor(month)+factor(MSA.Code),data=data_trans)
rob.ols_trans<-coeftest(ols_trans,vcov = vcovHC(ols_trans,type = "HC3"))
# rob.ols_trans
```
# 3.3 first-stage 2SLS regression
```{r}
st_trans1<-lm(onsite_pct~iv+reopened+net_mig_rate+covid_cases+vaccine+vrm_19pct+log(pop_million)+factor(month)+factor(MSA.Code),data=data_trans)

summary(st_trans1)$r.squared
st_trans1_pred<-st_trans1$fitted.values
data_trans$st1_pred<-st_trans1_pred

rob.st_trans1<-coeftest(st_trans1,vcov = vcovHC(st_trans1,type = "HC3"))
```
# 3.4 second-stage 2SLS regression
```{r}
# 3.4.1 main
st2_trans<-lm(ridership_19pct~st1_pred+reopened+net_mig_rate+covid_cases+vaccine+vrm_19pct+log(pop_million)+factor(month)+factor(MSA.Code),data=data_trans)
rob.st2_trans<-coeftest(st2_trans,vcov = vcovHC(st2_trans,type = "HC3"))

# 3.4.1 spatial heteogeneity
st2_trans_trans<-lm(ridership_19pct~st1_pred+I(st1_pred*log(transitcommute_pct))+reopened+net_mig_rate+covid_cases+vaccine+vrm_19pct+log(pop_million)+factor(month)+factor(MSA.Code),data=data_trans)

rob.st2_trans_trans<-coeftest(st2_trans_trans,vcov = vcovHC(st2_trans_trans,type = "HC3"))

# 3.4.2 time heteogeneity
st2_trans_time<-lm(ridership_19pct~I(st1_pred*((date>=as.Date('2020-04-01'))&(date<as.Date('2020-07-01'))))+I(st1_pred*((date>=as.Date('2020-07-01'))&(date<as.Date('2020-10-01'))))+I(st1_pred*((date>=as.Date('2020-10-01'))&(date<as.Date('2021-01-01'))))+I(st1_pred*((date>=as.Date('2021-01-01'))&(date<as.Date('2021-04-01'))))+I(st1_pred*((date>=as.Date('2021-04-01'))&(date<as.Date('2021-07-01'))))+I(st1_pred*((date>=as.Date('2021-07-01'))&(date<as.Date('2021-10-01'))))+I(st1_pred*((date>=as.Date('2021-10-01'))&(date<as.Date('2022-01-01'))))+I(st1_pred*((date>=as.Date('2022-01-01'))&(date<as.Date('2022-04-01'))))+reopened+net_mig_rate+covid_cases+vaccine+vrm_19pct+log(pop_million)+factor(month)+factor(MSA.Code),data=data_trans)

rob.st2_trans_time<-coeftest(st2_trans_time,vcov = vcovHC(st2_trans_time,type = "HC3"))
```

# 3.5 calculate the effects on transit fares
```{r}
fare<-read.csv('auxiliary_data/fare.csv')

results_trans<- tidy(rob.st2_trans_trans)
# results_trans

cov_inter<-cluster.vcov(st2_trans_trans, data_trans$MSA.Code)['st1_pred','I(st1_pred * log(transitcommute_pct))']
var_main<-cluster.vcov(st2_trans_trans, data_trans$MSA.Code)['st1_pred','st1_pred']
var_2<-cluster.vcov(st2_trans_trans, data_trans$MSA.Code)['I(st1_pred * log(transitcommute_pct))','I(st1_pred * log(transitcommute_pct))']

trans19_avg<-data_trans%>%
  filter(year==2021)%>%
  group_by(MSA.Code)%>%
  dplyr::summarise(ridership2019=mean(ridership2019))


df_effect_trans<-data_trans%>%
  filter(!is.na(MSA.Name))%>%
  filter(date==as.Date('2020-04-01'))%>%
  select(MSA.Title,MSA.Name,MSA.Code,transitcommute_pct)%>%
  left_join(trans19_avg,by='MSA.Code')%>%
  mutate(Coefficient=rob.st2_trans_trans['st1_pred','Estimate']+rob.st2_trans_trans['I(st1_pred * log(transitcommute_pct))','Estimate']*log(transitcommute_pct))%>%
  mutate(SE=sqrt(var_main+var_2*(log(transitcommute_pct)**2)+2*cov_inter*log(transitcommute_pct)))%>%
  mutate(Coef.low_95=Coefficient-1.96*SE,Coef.high_95=Coefficient+1.96*SE)%>%
  mutate(trans_effect=Coefficient*ridership2019/100,trans.low_95=Coef.low_95*ridership2019/100,trans.high_95=Coef.high_95*ridership2019/100)%>%
  left_join(fare,by='MSA.Code')%>%
  mutate(fare_effect=trans_effect*fare,fare.low_95=trans.low_95*fare,fare.high_95=trans.high_95*fare)

df_effect_trans2<-df_effect_trans %>%
  dplyr::filter(MSA.Name%in%df_effect_cong$MSA.Name)

df_effect_trans
```

```{r}
factor=100000
trans <- function(x) {
  ifelse(x > 2600000/factor, x/10+1040000/factor, x)
}
inv <- function(x) {
  ifelse(x > 1300000/factor, (x-1040000/factor)*10, x)
}
my_trans <- trans_new("my_trans", trans,inv)
```



# 3.5.1 plot the effects on transit ridership
```{r}
df_effect_trans2 %>%
  mutate(MSA.Name = fct_reorder(MSA.Name, trans_effect))%>%
  # filter(MSA.Name!='New York NY')%>%
ggplot(aes(x = MSA.Name, y = trans_effect/factor)) +
        geom_hline(yintercept = 0,
                   colour = gray(1/2), lty = 2) +
        geom_bar(stat="identity",fill="mediumpurple4") +
        geom_linerange(aes(x = MSA.Name, 
                     ymin = trans.low_95/factor,
                     ymax = trans.high_95/factor),
                   lwd = 1/2,position = position_dodge(width = 0),color='orange2')+
  geom_text(aes(label = round(trans_effect/factor,2)),hjust = -0.5,size = 7)+ scale_y_continuous(trans = my_trans,breaks = c(seq(0,1200000/factor,100000/factor),13,60,96))+
  coord_flip()+
    labs(y=bquote('Effect on monthly transit ridership (in 100,000)'))+
  labs(x="MSA")+
  theme(legend.key=element_blank(),
        axis.ticks=element_blank(),
        axis.text.x=element_text(size=20,family="Arial"), 
        axis.text.y=element_text(size=20,family="Arial"), 
        axis.title=element_text( size=25, family="Arial"),
        axis.title.y = element_text(margin = margin(r = 12,l=10)),
        axis.title.x = element_text(margin = margin(t = 10)),
        legend.text=element_text(size=20, family="Arial",margin = margin(t = 7)), 
        panel.background=element_blank(),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.grid.major.x = element_line(colour = "lightgrey"),
        panel.grid.major.y = element_blank(),
        # panel.grid.minor.x = element_line(colour = "lightgrey"),
        # panel.grid.minor.y = element_blank(),
        # panel.grid.minor = element_line(colour = "lightgrey"),
        legend.box.background = element_rect(colour = "black"),
        text = element_text(size = 20),
        plot.margin = margin(t = 0.5,  # Top margin
                             r = 0.5,  # Right margin
                             b = 0.2,  # Bottom margin
                             l = 0, "cm"))
# ggsave('plots/effect_trans.png', dpi = 800, width = 15, height = 16, units = "in")
```
# 3.5.2 plot the effects on transit fares
```{r}
df_effect_trans2 %>%
  mutate(MSA.Name = fct_reorder(MSA.Name, fare_effect))%>%
  # filter(MSA.Name!='New York NY')%>%
ggplot(aes(x = MSA.Name, y = fare_effect/factor)) +
        geom_hline(yintercept = 0,
                   colour = gray(1/2), lty = 2) +
        geom_bar(stat="identity",fill="mediumpurple4") +
        geom_linerange(aes(x = MSA.Name, 
                     ymin = fare.low_95/factor,
                     ymax = fare.high_95/factor),
                   lwd = 1/2,position = position_dodge(width = 0),color='orange2')+
  geom_text(aes(label = round(fare_effect/factor,2)),hjust = -0.5,size = 7)+ scale_y_continuous(trans = my_trans,breaks = c(seq(0,1800000/factor,200000/factor),48,111,174))+
  coord_flip()+
    labs(y=bquote('Effect on monthly transit fare revenue (in $100,000)'))+
  labs(x="MSA")+
  theme(legend.key=element_blank(),
        axis.ticks=element_blank(),
        axis.text.x=element_text(size=20,family="Arial"), 
        axis.text.y=element_text(size=20,family="Arial"), 
        axis.title=element_text( size=25, family="Arial"),
        axis.title.y = element_text(margin = margin(r = 12,l=10)),
        axis.title.x = element_text(margin = margin(t = 10)),
        legend.text=element_text(size=20, family="Arial",margin = margin(t = 7)), 
        panel.background=element_blank(),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.grid.major.x = element_line(colour = "lightgrey"),
        panel.grid.major.y = element_blank(),
        # panel.grid.minor.x = element_line(colour = "lightgrey"),
        # panel.grid.minor.y = element_blank(),
        # panel.grid.minor = element_line(colour = "lightgrey"),
        legend.box.background = element_rect(colour = "black"),
        text = element_text(size = 20),
        plot.margin = margin(t = 0.5,  # Top margin
                             r = 0.5,  # Right margin
                             b = 0.2,  # Bottom margin
                             l = 0, "cm"))
ggsave('plots/effect_trans_fare.png', dpi = 800, width = 15, height = 16, units = "in")
```
# 3.5.3 plot the model coefficient for each MSA
```{r}
df_effect_trans2 %>%
  mutate(MSA.Name = fct_reorder(MSA.Name, Coefficient))%>%
ggplot(aes(x = MSA.Name, y = Coefficient)) +
        geom_hline(yintercept = 0,
                   colour = gray(1/2), lty = 2) +
        geom_point(aes(x = MSA.Name,
                    y = Coefficient),position = position_dodge(width = 0),size=3) +
        geom_linerange(aes(x = MSA.Name,
                     ymin = Coef.low_95,
                     ymax = Coef.high_95),
                   lwd = 1,position = position_dodge(width = 0)) +
        coord_flip()+
  labs(x="MSA")+
  theme(legend.key=element_blank(),
        axis.ticks=element_blank(),
        axis.text.x=element_text(size=20,family="Arial"), 
        axis.text.y=element_text(size=20,family="Arial"), 
        axis.title=element_text( size=25, family="Arial"),
        axis.title.y = element_text(margin = margin(r = 12)),
        axis.title.x = element_text(margin = margin(t = 10)),
        legend.text=element_text(size=20, family="Arial",margin = margin(t = 7)), 
        panel.background=element_blank(),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.grid.major.x = element_line(colour = "lightgrey"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_line(colour = "lightgrey"),
        panel.grid.minor.y = element_blank(),
        # panel.grid.minor = element_line(colour = "lightgrey"),
        legend.box.background = element_rect(colour = "black"),
        text = element_text(size = 20),
        plot.margin = margin(t = 0.5,  # Top margin
                             r = 0.5,  # Right margin
                             b = 0.2,  # Bottom margin
                             l = 0.2, "cm"))
ggsave('plots/coef_trans.png', dpi = 800, width = 12, height = 16, units = "in")
```

# 4 Summarize modeling results
# 4.1 Summarize the second stage results
```{r}
stargazer(ols_vmt,st2_st_main,st2_st_trans,ols_cong,st2_cong,st2_cong_pop,ols_trans,st2_trans,st2_trans_trans,se = list(rob.ols_vmt[,"Std. Error"],rob.st2_st_main[,"Std. Error"],rob.st2_st_trans[,"Std. Error"],rob.ols_cong[,"Std. Error"],rob.st2_cong[,"Std. Error"],rob.st2_cong_pop[,"Std. Error"],rob.ols_trans[,"Std. Error"],rob.st2_trans[,"Std. Error"],rob.st2_trans_trans[,"Std. Error"]))
```
# 4.2 Summarize the first stage results
```{r}
stargazer(st_state1,st_cong1,st_trans1,se = list(rob.st_state1[,"Std. Error"],rob.st_cong1[,"Std. Error"],rob.st_trans1[,"Std. Error"]))
```

# 5. Plot time-varying effects
# 5.1 Summarize main effects by time
```{r}
library(zoo)
########### VMT
results_vmt <- tidy(rob.st2_st_time)
results_vmt2<-results_vmt%>%
           dplyr::rename(Variable = term,
                  Coefficient = estimate,
                  SE = std.error) %>% select(-statistic,
                              -p.value)%>%
  mutate(conf.low_95=Coefficient-1.645*SE,conf.high_95=Coefficient+1.645*SE)

from <- as.Date('2020-04-01')
to <- as.Date('2022-03-01')

results_vmt3<-results_vmt2[2:9,]
results_vmt3$yq<-seq(as.yearqtr(from), as.yearqtr(to), by = 1/4)
results_vmt3$type<-'VMT'

########### congestion
results_cong <- tidy(rob.st2_cong_time)
results_cong2<-results_cong%>%
           dplyr::rename(Variable = term,
                  Coefficient = estimate,
                  SE = std.error) %>% select(-statistic,
                              -p.value)%>%
  mutate(conf.low_95=Coefficient-1.645*SE,conf.high_95=Coefficient+1.645*SE)

# results_cong2_1<-results_cong2
# results_cong2_1[c('Coefficient','conf.low_95','conf.high_95')]<-results_cong2[c('Coefficient','conf.low_95','conf.high_95')]*30

from <- as.Date('2020-04-01')
to <- as.Date('2022-03-01')

results_cong3<-results_cong2[2:8,]
results_cong3[8,] <- NA
results_cong3$yq<-seq(as.yearqtr(from), as.yearqtr(to), by = 1/4)
results_cong3$type<-'Congestion hours'

########### transit
results_trans <- tidy(rob.st2_trans_time)
results_trans2<-results_trans%>%
           dplyr::rename(Variable = term,
                  Coefficient = estimate,
                  SE = std.error) %>% select(-statistic,
                              -p.value)%>%
  mutate(conf.low_95=Coefficient-1.645*SE,conf.high_95=Coefficient+1.645*SE)

from <- as.Date('2020-04-01')
to <- as.Date('2022-03-01')

results_trans3<-results_trans2[2:9,]
results_trans3$yq<-seq(as.yearqtr(from), as.yearqtr(to), by = 1/4)
results_trans3$type<-'Transit ridership'

results_time_full<-rbind(results_cong3,results_trans3,results_vmt3)

# results_trans2$Date<-seq(as.Date("2020/04/01"), by = "month", length.out = 24)
# iv_sel<-iv_total[2:25,]
results_time_full$type=factor(results_time_full$type,levels=c('Congestion hours', 'Transit ridership','VMT'))
```

```{r}
results_time_vmt_trans<-results_time_full%>%
  filter(type%in%c('VMT', 'Transit ridership'))
results_time_vmt_trans

results_time_cong<-results_time_full%>%
  filter(type%in%c('Congestion hours'))%>%
  filter(!is.na(Coefficient))
```
# 5.2 Plot time-varying effects
```{r}
## for VMT + transit
ggplot(data=results_time_vmt_trans,aes(x = yq, 
                    y = Coefficient,group=type,color=type,shape=type))+ geom_smooth(method=lm,size=0.7,linetype='dashed',se=F)+
        geom_point(position=position_dodge(width = 0.1),size=3,stroke=1) + 
          geom_line(position=position_dodge(width = 0.1),lwd = 0.7) +
        geom_linerange(aes(x = yq,
                     ymin = conf.low_95,
                     ymax = conf.high_95),alpha=0.7,
                   lwd = 0.5,position=position_dodge(width = 0.1))+
scale_color_manual('',values=c("#F8766D","#00BA38","#619CFF"),labels=c('VMT', 'Transit ridership','Congestion hours'),breaks=c('VMT', 'Transit ridership','Congestion hours'))+
  scale_shape_manual('',values = c(0, 1, 2),labels=c('VMT', 'Transit ridership','Congestion hours'),breaks=c('VMT', 'Transit ridership','Congestion hours'))+
   geom_hline(yintercept= 0,linetype='dashed')+
  # scale_y_continuous(name = "Marginal effects on VMT \nand transit ridership", sec.axis = sec_axis(~./30, name = "Marginal effect on log(CH)"))+
  zoo::scale_x_yearqtr(breaks = seq(from = min(results_time_full$yq), to = max(results_time_full$yq), by = 0.25),format = '%Y \nQ%q')+
  labs(x="")+
  theme(legend.position = c(0.6, 0.2))+
  theme(legend.key=element_blank(),
        # axis.ticks=element_blank(),
        axis.ticks.length=unit(.25, "cm"),
        axis.text=element_text(size=20,family="Arial"), 
        axis.text.x = element_text(vjust = 0.5, hjust=0.5),
        axis.title.y = element_text(size=25, family="Arial",margin = margin(r = 12)),
        axis.title.y.right = element_text(size=25, family="Arial",margin = margin(l = 12)),
        legend.box.background = element_rect(fill="transparent",colour=NA),
        legend.background = element_rect(fill='transparent'),
        legend.text=element_text(size=20, family="Arial",margin = margin(t = 3)), 
        panel.background=element_blank(),
        panel.border = element_rect(colour = 'black', fill=NA),
        panel.grid.major = element_line(colour = "lightgrey"), 
        panel.grid.minor = element_line(colour = "lightgrey"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        # legend.box.background = element_rect(colour = "black"),
        text = element_text(size = 20),
        plot.margin = margin(t = 0.5,  # Top margin
                             r = 1,  # Right margin
                             b = 0,  # Bottom margin
                             l = 0.5, "cm"))
ggsave('plots/effect_time_vmt_trans.png', dpi = 800, width = 7, height = 7, units = "in")
```

```{r}
## for congestion
ggplot(data=results_time_cong,aes(x = yq, 
                    y = Coefficient,group=type,color=type,shape=type))+ geom_smooth(method=lm,size=0.7,linetype='dashed',se=F)+
        geom_point(position=position_dodge(width = 0.1),size=3,stroke=1) + 
          geom_line(position=position_dodge(width = 0.1),lwd = 0.7) +
        geom_linerange(aes(x = yq,
                     ymin = conf.low_95,
                     ymax = conf.high_95),alpha=0.7,
                   lwd = 0.5,position=position_dodge(width = 0.1))+
scale_color_manual('',values=c("#F8766D","#00BA38","#619CFF"),labels=c('VMT', 'Transit ridership','Congestion hours'),breaks=c('VMT', 'Transit ridership','Congestion hours'))+
  scale_shape_manual('',values = c(0, 1, 2),labels=c('VMT', 'Transit ridership','Congestion hours'),breaks=c('VMT', 'Transit ridership','Congestion hours'))+
   geom_hline(yintercept= 0,linetype='dashed')+
  # scale_y_continuous(name = "Marginal effects on VMT \nand transit ridership", sec.axis = sec_axis(~./30, name = "Marginal effect on log(CH)"))+
  zoo::scale_x_yearqtr(breaks = seq(from = min(results_time_full$yq), to = max(results_time_full$yq), by = 0.25),format = '%Y \nQ%q')+
  labs(x="")+
  theme(legend.position = c(0.6, 0.2))+
  theme(legend.key=element_blank(),
        # axis.ticks=element_blank(),
        axis.ticks.length=unit(.25, "cm"),
        axis.text=element_text(size=20,family="Arial"), 
        axis.text.x = element_text(vjust = 0.5, hjust=0.5),
        axis.title.y = element_text(size=25, family="Arial",margin = margin(r = 12)),
        axis.title.y.right = element_text(size=25, family="Arial",margin = margin(l = 12)),
        legend.box.background = element_rect(fill="transparent",colour=NA),
        legend.background = element_rect(fill='transparent'),
        legend.text=element_text(size=20, family="Arial",margin = margin(t = 3)), 
        panel.background=element_blank(),
        panel.border = element_rect(colour = 'black', fill=NA),
        panel.grid.major = element_line(colour = "lightgrey"), 
        panel.grid.minor = element_line(colour = "lightgrey"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        # legend.box.background = element_rect(colour = "black"),
        text = element_text(size = 20),
        plot.margin = margin(t = 0.5,  # Top margin
                             r = 1,  # Right margin
                             b = 0,  # Bottom margin
                             l = 0.5, "cm"))
ggsave('plots/effect_time_cong.png', dpi = 800, width = 7, height = 7, units = "in")
```

