---
title: "01models"
output: html_document
---

```{r setup, include=FALSE}
library(miceadds)
library(lmtest)
library(multiwayvcov)
library(modelsummary)
library(fabricatr)
library(mediation)
library(ivpack)
library(tidyr)
library(lubridate)
library(dplyr)
library(stargazer)
library(plm)
library(lfe)
library(ggplot2)
library(forcats)
library(scales)
library(ivmodel)
library(xtable)
```

```{r}
#'gdp_pct': state gdp compared to the same month in 2019
```

# 1. VMT model
# 1.1 load data
```{r}
begin_date='2020-03-01'
end_date='2022-11-01'
data_vmt0<-read.csv('data_processed/data_vmt.csv')
data_vmt0$date<-as.Date(data_vmt0$date)
data_vmt0$month<-month(data_vmt0$date)
data_vmt0$year<-year(data_vmt0$date)

st_region_map<-read.csv('auxiliary_data/st_region_map.csv')
st_region_map<-st_region_map%>%
  dplyr::select(State.Code,Region,Division)%>%
  dplyr::rename(state=State.Code)

data_vmt<-data_vmt0%>%
  filter((date>as.Date(begin_date))&(date<as.Date(end_date)))%>%
  left_join(st_region_map,by='state')
data_vmt
```


```{r}
# See transitcommute_pct by states
data_vmt%>%
  group_by(Division)%>%
  dplyr::summarise(log_transitcommute_pct=mean(log(transitcommute_pct+1)),
                   transitcommute_pct=mean(transitcommute_pct))%>%
  arrange(desc(transitcommute_pct))
```


# 1.2 fixed-effects regression
```{r}

ols_vmt<-lm(vmt_19pct~onsite_pct+reopened+gdp_p_in1000+UR+vrm_19pct+fare+cases_p_1000+dose_p_lag3+dose_p_lag3_6+dose_p_lag6_9+net_mig_rate+l_pop_million+factor(state)+factor(month),data=data_vmt)

rob.ols_vmt<-coeftest(ols_vmt,vcov = vcovHC(ols_vmt,type = "HC3"))

########
ols_vmt_nofix<-lm(vmt_19pct~onsite_pct+reopened+gdp_p_in1000+UR+vrm_19pct+fare+cases_p_1000+dose_p_lag3+dose_p_lag3_6+dose_p_lag6_9+net_mig_rate+l_pop_million,data=data_vmt)
rob.ols_vmt_nofix<-coeftest(ols_vmt_nofix,vcov = vcovHC(ols_vmt_nofix,type = "HC3"))




```



# 1.3 first-stage 2SLS regression
```{r}
st_state1<-lm(onsite_pct~iv+reopened+gdp_p_in1000+UR+vrm_19pct+fare+cases_p_1000+dose_p_lag3+dose_p_lag3_6+dose_p_lag6_9+net_mig_rate+l_pop_million+factor(state)+factor(month),data=data_vmt)
rob.st_state1<-coeftest(st_state1,vcov = vcovHC(st_state1,type = "HC3"))

summary(st_state1)$r.squared
st_state1_pred<-st_state1$fitted.values
data_vmt$st1_pred<-st_state1_pred
```



# 1.4 second-stage 2SLS regression
```{r}
# 1.4.1 main
st2_st_main<-lm(vmt_19pct~st1_pred+reopened+gdp_p_in1000+UR+vrm_19pct+fare+cases_p_1000+dose_p_lag3+dose_p_lag3_6+dose_p_lag6_9+net_mig_rate+l_pop_million+factor(state)+factor(month),data=data_vmt)

rob.st2_st_main<-coeftest(st2_st_main,vcov = vcovHC(st2_st_main,type = "HC3"))

# 1.4.2 spatial heteogeneity
st2_st_trans<-lm(vmt_19pct~st1_pred+I(st1_pred*log(transitcommute_pct))+reopened+gdp_p_in1000+UR+vrm_19pct+fare+cases_p_1000+dose_p_lag3+dose_p_lag3_6+dose_p_lag6_9+net_mig_rate+l_pop_million+factor(state)+factor(month),data=data_vmt)

rob.st2_st_trans<-coeftest(st2_st_trans,vcov = vcovHC(st2_st_trans,type = "HC3"))

st2_st_region<-lm(vmt_19pct~st1_pred:Division+reopened+gdp_p_in1000+UR+vrm_19pct+fare+cases_p_1000+dose_p_lag3+dose_p_lag3_6+dose_p_lag6_9+net_mig_rate+l_pop_million+factor(state)+factor(month),data=data_vmt)

rob.st2_st_region<-coeftest(st2_st_region,vcov = vcovHC(st2_st_region,type = "HC3"))

# 1.4.3 time heteogeneity
st2_st_time<-lm(vmt_19pct~I(st1_pred*((date>=as.Date('2020-04-01'))&(date<as.Date('2020-07-01'))))+I(st1_pred*((date>=as.Date('2020-07-01'))&(date<as.Date('2020-10-01'))))+I(st1_pred*((date>=as.Date('2020-10-01'))&(date<as.Date('2021-01-01'))))+I(st1_pred*((date>=as.Date('2021-01-01'))&(date<as.Date('2021-04-01'))))+I(st1_pred*((date>=as.Date('2021-04-01'))&(date<as.Date('2021-07-01'))))+I(st1_pred*((date>=as.Date('2021-07-01'))&(date<as.Date('2021-10-01'))))+I(st1_pred*((date>=as.Date('2021-10-01'))&(date<as.Date('2022-01-01'))))+I(st1_pred*((date>=as.Date('2022-01-01'))&(date<as.Date('2022-04-01'))))+I(st1_pred*((date>=as.Date('2022-04-01'))&(date<as.Date('2022-07-01'))))+I(st1_pred*((date>=as.Date('2022-07-01'))&(date<=as.Date('2022-10-01'))))+reopened+gdp_p_in1000+UR+vrm_19pct+fare+cases_p_1000+dose_p_lag3+dose_p_lag3_6+dose_p_lag6_9+net_mig_rate+l_pop_million+factor(state)+factor(month),data=data_vmt)

rob.st2_st_time<-coeftest(st2_st_time,vcov = vcovHC(st2_st_time,type = "HC3"))

```

# 1.5 calculate CO2 emissions
# 1.5.1 load VMT-CO2 data
```{r}
co2_df<-read.csv('auxiliary_data/co2_lbs_mile2120.csv')
co2_df<-co2_df%>%
  dplyr::select(state,co2_lbs_mile)
```
# 1.5.2 calculate CO2 effect for each state
```{r}
results_st <- tidy(rob.st2_st_trans)
results_st

cov_inter<-cluster.vcov(st2_st_trans, data_vmt$state)['st1_pred','I(st1_pred * log(transitcommute_pct))']
var_main<-cluster.vcov(st2_st_trans, data_vmt$state)['st1_pred','st1_pred']
var_2<-cluster.vcov(st2_st_trans, data_vmt$state)['I(st1_pred * log(transitcommute_pct))','I(st1_pred * log(transitcommute_pct))']

vmt19_avg<-data_vmt%>%
  filter(year==2021)%>%
  group_by(state)%>%
  dplyr::summarise(vmt19=mean(vmt19))
```
```{r}
df_effect<-data_vmt%>%
  dplyr::select(state,transitcommute_pct,iv,pop_million)%>%
  dplyr::group_by(state)%>%
  dplyr::summarise(transitcommute_pct=mean(transitcommute_pct),iv=mean(iv),pop_million=mean(pop_million))%>%
  dplyr::left_join(vmt19_avg,by='state')%>%
  dplyr::mutate(l_trans=log(transitcommute_pct))%>%
  dplyr::mutate(Coefficient=rob.st2_st_trans['st1_pred','Estimate']+rob.st2_st_trans['I(st1_pred * log(transitcommute_pct))','Estimate']*log(transitcommute_pct))%>%
  dplyr::mutate(SE=sqrt(var_main+var_2*(log(transitcommute_pct)**2)+2*cov_inter*log(transitcommute_pct)))%>%
  dplyr::mutate(Coef.low_95=Coefficient-1.96*SE,Coef.high_95=Coefficient+1.96*SE)%>%
  dplyr::mutate(vmt_effect_mill=Coefficient*vmt19/100,vmt.low_95=Coef.low_95*vmt19/100,vmt.high_95=Coef.high_95*vmt19/100)%>%
  dplyr::mutate(vmt_p_effect=vmt_effect_mill/pop_million,vmt_p.low_95=vmt.low_95/pop_million,vmt_p.high_95=vmt.high_95/pop_million)%>% #unit: miles
  dplyr::left_join(co2_df,by='state')%>%
  dplyr::mutate(co2_effect_mill_lbs=vmt_effect_mill*co2_lbs_mile,co2.low_95=vmt.low_95*co2_lbs_mile,co2.high_95=vmt.high_95*co2_lbs_mile)%>%
  dplyr::mutate(co2_p=vmt_p_effect*co2_lbs_mile,co2_p.low_95=vmt_p.low_95*co2_lbs_mile,co2_p.high_95=vmt_p.high_95*co2_lbs_mile)%>%
  dplyr::arrange(desc(co2_p))

df_effect<-df_effect%>%
  mutate(co2_effect_thous_tons=co2_effect_mill_lbs*1000/2204.62)%>%
  mutate(co2.low_95_thous_tons=co2.low_95*1000/2204.62,co2.high_95_thous_tons=co2.high_95*1000/2204.62)%>% 
  mutate(co2_effect_mill_lbs_yr=co2_effect_mill_lbs*12)%>%
  mutate(co2.low_95_mill_lbs_yr=co2.low_95*12,co2.high_95_mill_lbs_yr=co2.high_95*12)%>%
  mutate(co2_effect_mill_tons_yr=co2_effect_mill_lbs_yr/2204.62)%>%
  mutate(co2.low_95_mill_tons_yr=co2.low_95_mill_lbs_yr/2204.62,co2.high_95_mill_tons_yr=co2.high_95_mill_lbs_yr/2204.62)%>%
  mutate(co2_p_yr=co2_p*12)%>%
  mutate(co2_p.low_95_yr=co2_p.low_95*12,co2_p.high_95_yr=co2_p.high_95*12)

```

# 1.5.3 plot CO2 effect for each state
```{r}
df_effect %>%
  mutate(state = fct_reorder(state, co2_effect_thous_tons))%>%
ggplot(aes(x = state, y = co2_effect_thous_tons)) +
        geom_hline(yintercept = 0,
                   colour = gray(1/2), lty = 2) +
        geom_bar(stat="identity",fill='steelblue4') +
        geom_linerange(aes(x = state,
                     ymin = co2.low_95_thous_tons,
                     ymax = co2.high_95_thous_tons),
                   lwd = 0.7,position = position_dodge(width = 0),color='orange2')+
  geom_text(aes(label = round(co2_effect_thous_tons,1)),hjust = -0.5,size = 7) +
        coord_flip()+
    labs(y=bquote('Effect on monthly' ~CO[2]*' emissions (thousand metric tons)'))+
  labs(x="State")+
  theme(legend.key=element_blank(),
        axis.ticks=element_blank(),
        axis.text.x=element_text(size=20,family="Arial"),
        axis.text.y=element_text(size=20,family="Arial"),
        axis.title=element_text( size=25, family="Arial"),
        axis.title.y = element_text(margin = margin(r = 12)),
        axis.title.x = element_text(margin = margin(t = 10)),
        legend.text=element_text(size=20, family="Arial",margin = margin(t = 7)),
        panel.background=element_blank(),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.grid.major.x = element_line(colour = "lightgrey"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_line(colour = "lightgrey"),
        panel.grid.minor.y = element_blank(),
        # panel.grid.minor = element_line(colour = "lightgrey"),
        legend.box.background = element_rect(colour = "black"),
        text = element_text(size = 20),
        plot.margin = margin(t = 0.5,  # Top margin
                             r = 0.5,  # Right margin
                             b = 0.2,  # Bottom margin
                             l = 0.2, "cm"))
# ggsave('plots/co2_vmt.png', dpi = 800, width = 12, height = 16, units = "in")
```
# 1.5.4 calculate the aggregate CO2 effect
```{r}
total_us_co2=1939.46-11.9-12.3

##VMT-related CO2 (yearly)
vmt_co2_19=sum(df_effect$vmt19*df_effect$co2_lbs_mile)*12

##use official transport-sector emission as the baseline
co2_sum=sum(df_effect$co2_effect_mill_lbs)*12/2204.62 #MtCO2

print(paste0('1% increase in remote work will cause reduction in annual CO2 emission: ',round(co2_sum,2),' million metric tons, ',round(co2_sum*100/total_us_co2,2),'% of the 2019 level'))
```

# 1.6 plot the coefficient for each state
```{r}
df_effect %>%
  mutate(state = fct_reorder(state, Coefficient))%>%
ggplot(aes(x = state, y = Coefficient)) +
        geom_hline(yintercept = 0,
                   colour = gray(1/2), lty = 2) +
        geom_point(aes(x = state,
                    y = Coefficient),position = position_dodge(width = 0),size=3) +
        geom_linerange(aes(x = state,
                     ymin = Coef.low_95,
                     ymax = Coef.high_95),
                   lwd = 1,position = position_dodge(width = 0)) +
        coord_flip()+
  labs(x="State")+
  theme(legend.key=element_blank(),
        axis.ticks=element_blank(),
        axis.text.x=element_text(size=20,family="Arial"), 
        axis.text.y=element_text(size=20,family="Arial"), 
        axis.title=element_text( size=25, family="Arial"),
        axis.title.y = element_text(margin = margin(r = 12)),
        axis.title.x = element_text(margin = margin(t = 10)),
        legend.text=element_text(size=20, family="Arial",margin = margin(t = 7)), 
        panel.background=element_blank(),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.grid.major.x = element_line(colour = "lightgrey"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_line(colour = "lightgrey"),
        panel.grid.minor.y = element_blank(),
        # panel.grid.minor = element_line(colour = "lightgrey"),
        legend.box.background = element_rect(colour = "black"),
        text = element_text(size = 20),
        plot.margin = margin(t = 0.5,  # Top margin
                             r = 0.5,  # Right margin
                             b = 0.2,  # Bottom margin
                             l = 0.2, "cm"))
# ggsave('plots/coef_vmt.png', dpi = 800, width = 12, height = 16, units = "in")
```


# 2. Transit model
# 2.1 load data
```{r}
end_date<-'2022-11-01'
data_trans0<-read.csv('data_processed/data_transit.csv')
data_trans0$date<-as.Date(data_trans0$date)
data_trans0$month<-month(data_trans0$date)
data_trans0$year<-year(data_trans0$date)
data_trans<-data_trans0%>%
  filter((date>as.Date('2020-03-01'))&(date<as.Date(end_date)))%>%
  filter(!is.na(onsite_pct))%>%
  left_join(st_region_map,by='state')

```


# 2.2 fixed-effects regression
```{r}
ols_trans<-lm(ridership_19pct~onsite_pct+reopened+gdp_p_in1000+UR+vrm_19pct+fare+cases_p_1000+dose_p_lag3+dose_p_lag3_6+dose_p_lag6_9+net_mig_rate+l_pop_million+factor(MSA.Code)+factor(month),data=data_trans)

rob.ols_trans<-coeftest(ols_trans,vcov = vcovHC(ols_trans,type = "HC3"))
rob.ols_trans

ols_trans_nofix<-lm(ridership_19pct~onsite_pct+reopened+gdp_p_in1000+UR+vrm_19pct+fare+cases_p_1000+dose_p_lag3+dose_p_lag3_6+dose_p_lag6_9+net_mig_rate+l_pop_million,data=data_trans)
summary(ols_trans_nofix)
rob.ols_trans_nofix<-coeftest(ols_trans_nofix,vcov = vcovHC(ols_trans_nofix,type = "HC3"))
rob.ols_trans_nofix

```



# 2.3 first-stage 2SLS regression
```{r}
st_trans1<-lm(onsite_pct~iv+reopened+gdp_p_in1000+UR+vrm_19pct+fare+cases_p_1000+dose_p_lag3+dose_p_lag3_6+dose_p_lag6_9+net_mig_rate+l_pop_million+factor(MSA.Code)+factor(month),data=data_trans)

summary(st_trans1)$r.squared
st_trans1_pred<-st_trans1$fitted.values
data_trans$st1_pred<-st_trans1_pred

rob.st_trans1<-coeftest(st_trans1,vcov = vcovHC(st_trans1,type = "HC3"))
```


# 2.4 second-stage 2SLS regression
```{r}
# 2.4.1 main
st2_trans<-lm(ridership_19pct~st1_pred+reopened+gdp_p_in1000+UR+vrm_19pct+fare+cases_p_1000+dose_p_lag3+dose_p_lag3_6+dose_p_lag6_9+net_mig_rate+l_pop_million+factor(MSA.Code)+factor(month),data=data_trans)
rob.st2_trans<-coeftest(st2_trans,vcov = vcovHC(st2_trans,type = "HC3"))



# 2.4.1 spatial heteogeneity
st2_trans_trans<-lm(ridership_19pct~st1_pred+I(st1_pred*log(transitcommute_pct))+reopened+gdp_p_in1000+UR+vrm_19pct+fare+cases_p_1000+dose_p_lag3+dose_p_lag3_6+dose_p_lag6_9+net_mig_rate+l_pop_million+factor(MSA.Code)+factor(month),data=data_trans)

rob.st2_trans_trans<-coeftest(st2_trans_trans,vcov = vcovHC(st2_trans_trans,type = "HC3"))

st2_trans_region<-lm(ridership_19pct~st1_pred:Division+reopened+gdp_p_in1000+UR+vrm_19pct+fare+cases_p_1000+dose_p_lag3+dose_p_lag3_6+dose_p_lag6_9+net_mig_rate+l_pop_million+factor(MSA.Code)+factor(month),data=data_trans)

rob.st2_trans_region<-coeftest(st2_trans_region,vcov = vcovHC(st2_trans_region,type = "HC3"))

# 2.4.2 time heteogeneity
st2_trans_time<-lm(ridership_19pct~I(st1_pred*((date>=as.Date('2020-04-01'))&(date<as.Date('2020-07-01'))))+I(st1_pred*((date>=as.Date('2020-07-01'))&(date<as.Date('2020-10-01'))))+I(st1_pred*((date>=as.Date('2020-10-01'))&(date<as.Date('2021-01-01'))))+I(st1_pred*((date>=as.Date('2021-01-01'))&(date<as.Date('2021-04-01'))))+I(st1_pred*((date>=as.Date('2021-04-01'))&(date<as.Date('2021-07-01'))))+I(st1_pred*((date>=as.Date('2021-07-01'))&(date<as.Date('2021-10-01'))))+I(st1_pred*((date>=as.Date('2021-10-01'))&(date<as.Date('2022-01-01'))))+I(st1_pred*((date>=as.Date('2022-01-01'))&(date<as.Date('2022-04-01'))))+I(st1_pred*((date>=as.Date('2022-04-01'))&(date<as.Date('2022-07-01'))))+I(st1_pred*((date>=as.Date('2022-07-01'))&(date<=as.Date('2022-10-01'))))+reopened+gdp_p_in1000+UR+vrm_19pct+fare+cases_p_1000+dose_p_lag3+dose_p_lag3_6+dose_p_lag6_9+net_mig_rate+l_pop_million+factor(MSA.Code)+factor(month),data=data_trans)

rob.st2_trans_time<-coeftest(st2_trans_time,vcov = vcovHC(st2_trans_time,type = "HC3"))

```
```{r}
top_msa_df<-read.csv('data_processed/top_msa_df.csv')

data_trans2<-data_trans%>%
  left_join(top_msa_df,by='MSA.Code')

```



# 2.5 calculate the effects on transit fares
```{r}
fare<-read.csv('auxiliary_data/fare.csv')

results_trans<- tidy(rob.st2_trans_trans)

cov_inter<-cluster.vcov(st2_trans_trans, data_trans$MSA.Code)['st1_pred','I(st1_pred * log(transitcommute_pct))']
var_main<-cluster.vcov(st2_trans_trans, data_trans$MSA.Code)['st1_pred','st1_pred']
var_2<-cluster.vcov(st2_trans_trans, data_trans$MSA.Code)['I(st1_pred * log(transitcommute_pct))','I(st1_pred * log(transitcommute_pct))']

trans19_avg<-data_trans%>%
  filter(year==2021)%>%
  group_by(MSA.Code)%>%
  dplyr::summarise(ridership2019=mean(ridership2019))%>%
  left_join(fare,by='MSA.Code')%>%
  dplyr::mutate(fare_rev19=ridership2019*fare)%>%
  dplyr::select(-fare)


df_effect_trans<-data_trans2%>%
  # filter(!is.na(MSA.Name))%>%
  filter(date==as.Date('2020-04-01'))%>%
  dplyr::select(MSA.Title,MSA.Name,MSA.Code,transitcommute_pct)%>%
  dplyr::left_join(trans19_avg,by='MSA.Code')%>%
  dplyr::mutate(Coefficient=rob.st2_trans_trans['st1_pred','Estimate']+rob.st2_trans_trans['I(st1_pred * log(transitcommute_pct))','Estimate']*log(transitcommute_pct))%>%
  dplyr::mutate(SE=sqrt(var_main+var_2*(log(transitcommute_pct)**2)+2*cov_inter*log(transitcommute_pct)))%>%
  dplyr::mutate(Coef.low_95=Coefficient-1.96*SE,Coef.high_95=Coefficient+1.96*SE)%>%
  dplyr::mutate(trans_effect=Coefficient*ridership2019/100,trans.low_95=Coef.low_95*ridership2019/100,trans.high_95=Coef.high_95*ridership2019/100)%>%
  dplyr::left_join(fare,by='MSA.Code')%>%
  dplyr::mutate(fare_effect=trans_effect*fare,fare.low_95=trans.low_95*fare,fare.high_95=trans.high_95*fare)

df_effect_trans2<-df_effect_trans %>%
  dplyr::filter(!is.na(MSA.Name))

df_effect_trans2
```


```{r}
factor=100000
df_plt<-df_effect_trans2%>%
  arrange(desc(trans_effect))
first_low_bound<-round(df_plt[1,'fare.low_95']/factor)-1
scale_intercept<-round(df_plt[2,'fare.high_95']/factor+3-first_low_bound/10)

trans <- function(x) {
  ifelse(x > first_low_bound, x/10+scale_intercept, x)
}
inv <- function(x) {
  ifelse(x > first_low_bound/10+scale_intercept, (x-scale_intercept)*10, x)
}
my_trans <- trans_new("my_trans", trans,inv)
```


# 2.5.2 plot the effects on transit fares
```{r}
df_effect_trans2 %>%
  mutate(MSA.Name = fct_reorder(MSA.Name, fare_effect))%>%
  # filter(MSA.Name!='New York NY')%>%
ggplot(aes(x = MSA.Name, y = fare_effect/factor)) +
        geom_hline(yintercept = 0,
                   colour = gray(1/2), lty = 2) +
        geom_bar(stat="identity",fill="mediumpurple4") +
        geom_linerange(aes(x = MSA.Name, 
                     ymin = fare.low_95/factor,
                     ymax = fare.high_95/factor),
                   lwd = 1/2,position = position_dodge(width = 0),color='orange2')+
  geom_text(aes(label = round(fare_effect/factor,2)),hjust = -0.5,size = 7)+ scale_y_continuous(trans = my_trans,breaks = c(seq(0,round(df_plt[2,'fare.high_95']/factor)+1,2),first_low_bound+1,round(df_plt[1,'fare_effect']/factor),round(df_plt[1,'fare.high_95']/factor)))+
  coord_flip()+
    labs(y=bquote('Effect on monthly transit fare revenue (in $100,000)'))+
  labs(x="MSA")+
  theme(legend.key=element_blank(),
        axis.ticks=element_blank(),
        axis.text.x=element_text(size=20,family="Arial"), 
        axis.text.y=element_text(size=20,family="Arial"), 
        axis.title=element_text( size=25, family="Arial"),
        axis.title.y = element_text(margin = margin(r = 12,l=10)),
        axis.title.x = element_text(margin = margin(t = 10)),
        legend.text=element_text(size=20, family="Arial",margin = margin(t = 7)), 
        panel.background=element_blank(),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.grid.major.x = element_line(colour = "lightgrey"),
        panel.grid.major.y = element_blank(),
        legend.box.background = element_rect(colour = "black"),
        text = element_text(size = 20),
        plot.margin = margin(t = 0.5,  # Top margin
                             r = 0.5,  # Right margin
                             b = 0.2,  # Bottom margin
                             l = 0, "cm"))
# ggsave('plots/effect_trans_fare.png', dpi = 800, width = 15, height = 16, units = "in")
```


```{r}
rider_loss=round(sum(df_effect_trans$trans_effect)*12/1000000000,2)
fare_loss=round(sum(df_effect_trans$fare_effect)*12/1000000000,2)
rider_19=sum(trans19_avg$ridership2019)*12/1000000000 #billion
fare_19=sum(trans19_avg$fare_rev19)*12/1000000000 #billion

print(paste0('1% increase in remote work will cause annual transit ridership loss: ',rider_loss,' billion,',round(rider_loss*100/rider_19,2),'%'))

print(paste0('1% increase in remote work will cause annual transit fare revenue loss: ',fare_loss,' billion,',round(fare_loss*100/fare_19,2),'%'))

ny_fare_loss=round(df_effect_trans2[df_effect_trans2$MSA.Name=='New York NY','fare_effect']*12/1000000000,2)
ny_fare_month_loss=round(df_effect_trans2[df_effect_trans2$MSA.Name=='New York NY','fare_effect']/1000000,2)
print(paste0('1% increase in remote work will cause NY fare revenue loss (annual): ',ny_fare_loss,' billion, which is the % of total: ',round(ny_fare_loss*100/fare_loss,2),'%'))
print(paste0('1% increase in remote work will cause NY fare revenue loss (monthly): ',ny_fare_month_loss,' million, which is the % of total: ',round(ny_fare_loss*100/fare_loss,2),'%'))
```


# 2.5.3 plot the model coefficient for each MSA
```{r}
df_effect_trans2 %>%
  mutate(MSA.Name = fct_reorder(MSA.Name, Coefficient))%>%
ggplot(aes(x = MSA.Name, y = Coefficient)) +
        geom_hline(yintercept = 0,
                   colour = gray(1/2), lty = 2) +
        geom_point(aes(x = MSA.Name,
                    y = Coefficient),position = position_dodge(width = 0),size=3) +
        geom_linerange(aes(x = MSA.Name,
                     ymin = Coef.low_95,
                     ymax = Coef.high_95),
                   lwd = 1,position = position_dodge(width = 0)) +
        coord_flip()+
  labs(x="MSA")+
  theme(legend.key=element_blank(),
        axis.ticks=element_blank(),
        axis.text.x=element_text(size=20,family="Arial"), 
        axis.text.y=element_text(size=20,family="Arial"), 
        axis.title=element_text( size=25, family="Arial"),
        axis.title.y = element_text(margin = margin(r = 12)),
        axis.title.x = element_text(margin = margin(t = 10)),
        legend.text=element_text(size=20, family="Arial",margin = margin(t = 7)), 
        panel.background=element_blank(),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.grid.major.x = element_line(colour = "lightgrey"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_line(colour = "lightgrey"),
        panel.grid.minor.y = element_blank(),
        # panel.grid.minor = element_line(colour = "lightgrey"),
        legend.box.background = element_rect(colour = "black"),
        text = element_text(size = 20),
        plot.margin = margin(t = 0.5,  # Top margin
                             r = 0.5,  # Right margin
                             b = 0.2,  # Bottom margin
                             l = 0.2, "cm"))
ggsave('plots/coef_trans.png', dpi = 800, width = 12, height = 16, units = "in")
```

# 3 Summarize modeling results

# 3.1 Summarize OLS results
```{r}
variable_mapping <- c(
  "l_val_upt" = "ln (monthly transit ridership)",
  "ridership_19pct" = "Transit ridership compared to the 2019 level (%)",
  "onsite_pct" = "Recovery rate of onsite workers",
  "iv" = "Percentage of suitable remote workers",
  "net_mig_rate" = "Net migration rate",
  "l_val_vrm" = "ln (monthly transit VRM)",
  "vrm_19pct" = "Transit service recovery rate",
  "fare" = "Transit fare",
  "cases_p_1000" = "COVID cases per 1000 people",
  "vaccine_p_1000" = "Vaccinations per 1000 people",
  "dose_p_lag3" = "Vaccinations per person in the past 3 months",
  "dose_p_lag3_6" = "Vaccinations per person over the past 3-6 months",
  "dose_p_lag6_9" = "Vaccinations per person over the past 6-9 months",
  "vaccine_p_1000" = "Vaccinations per 1000 person",
  "reopened" = "Reopening status",
  "pop_million" = "Population (in millions)",
  "l_pop_million" = "ln (population in millions)",
  "transitcommute_pct" = "Percentage of transit commuters (%)",
  "gdp_p_in1000" = "GDP per capita (in thousand dollars)",
  "gdp_pct" = "Per capita GDP compared to the 2019 level (%)",
  "inc_p_in1000" = "Annual per capita income (in thousand dollars)",
  "inc_p_pct" = "Per capita income compared to the 2019 level (%)",
  "UR" = "Unemployment rate",
  "st1_pred" = "Recovery rate of onsite workers",
  "I(st1_pred*log(transitcommute_pct))" = "Recovery rate of onsite workers x log (percentage of transit commuters)"
)
var_list <- c("onsite_pct","reopened", "gdp_p_in1000","UR", "vrm_19pct", "fare", "cases_p_1000", "dose_p_lag3", "dose_p_lag3_6","dose_p_lag6_9", "net_mig_rate", "l_pop_million")
var_list_new <- variable_mapping[var_list]

stargazer(ols_vmt_nofix,ols_vmt,ols_trans_nofix,ols_trans,se = list(rob.ols_vmt_nofix[,"Std. Error"],rob.ols_vmt[,"Std. Error"],rob.ols_trans_nofix[,"Std. Error"],rob.ols_trans[,"Std. Error"]), covariate.labels = var_list_new)
```

# 3.2 Summarize the second stage results
```{r}
var_list <- c("onsite_pct","I(st1_pred*log(transitcommute_pct))", "reopened", "gdp_p_in1000","UR", "vrm_19pct", "fare", "cases_p_1000", "dose_p_lag3", "dose_p_lag3_6","dose_p_lag6_9", "net_mig_rate", "l_pop_million")
var_list_new <- variable_mapping[var_list]

stargazer(st2_st_main,st2_st_trans,st2_st_region,st2_trans,st2_trans_trans,st2_trans_region,se = list(rob.st2_st_main[,"Std. Error"],rob.st2_st_trans[,"Std. Error"],rob.st2_st_region[,"Std. Error"],rob.st2_trans[,"Std. Error"],rob.st2_trans_trans[,"Std. Error"],rob.st2_trans_region[,"Std. Error"]), covariate.labels = var_list_new)
```

# 3.3 Summarize the first stage results
```{r}
var_list <- c("iv", "reopened", "gdp_p_in1000","UR", "vrm_19pct", "fare", "cases_p_1000", "dose_p_lag3", "dose_p_lag3_6","dose_p_lag6_9", "net_mig_rate", "l_pop_million")
var_list_new <- variable_mapping[var_list]
stargazer(st_state1,st_trans1,se = list(rob.st_state1[,"Std. Error"],rob.st_trans1[,"Std. Error"]), covariate.labels = var_list_new)
```

# 3.4 Summarize time-varying effects
```{r}
stargazer(st2_st_time,st2_trans_time,rob.st2_st_time[,"Std. Error"],rob.st2_trans_time[,"Std. Error"])
```






#SWAA result
```{r}

data_vmt_swaa<-data_vmt%>%
  filter(!is.na(swaa_wfh))%>%
  filter(swaa_sample_count>=10)

ols_vmt_swaa<-lm(vmt_19pct~swaa_wfh+reopened+gdp_p_in1000+UR+vrm_19pct+fare+cases_p_1000+dose_p_lag3+dose_p_lag3_6+dose_p_lag6_9+net_mig_rate+l_pop_million+factor(state)+factor(month),data=data_vmt_swaa)

rob.ols_vmt_swaa<-coeftest(ols_vmt_swaa,vcov = vcovHC(ols_vmt_swaa,type = "HC"))

st_state1_swaa<-lm(swaa_wfh~iv+reopened+gdp_p_in1000+UR+vrm_19pct+fare+cases_p_1000+dose_p_lag3+dose_p_lag3_6+dose_p_lag6_9+net_mig_rate+l_pop_million+factor(state)+factor(month),data=data_vmt_swaa)

rob.st_state1_swaa<-coeftest(st_state1_swaa,vcov = vcovHC(st_state1_swaa,type = "HC"))

summary(st_state1_swaa)$r.squared
st_state1_swaa_pred<-st_state1_swaa$fitted.values
data_vmt_swaa$st1_swaa_pred<-st_state1_swaa_pred

st2_st_main_swaa<-lm(vmt_19pct~st1_swaa_pred+reopened+gdp_p_in1000+UR+vrm_19pct+fare+cases_p_1000+dose_p_lag3+dose_p_lag3_6+dose_p_lag6_9+net_mig_rate+l_pop_million+factor(state)+factor(month),data=data_vmt_swaa)


rob.st2_st_main_swaa<-coeftest(st2_st_main_swaa,vcov = vcovHC(st2_st_main_swaa,type = "HC"))

st2_st_main_swaa_trans<-lm(vmt_19pct~st1_swaa_pred+I(st1_swaa_pred*log(transitcommute_pct))+reopened+gdp_p_in1000+UR+vrm_19pct+fare+cases_p_1000+dose_p_lag3+dose_p_lag3_6+dose_p_lag6_9+net_mig_rate+l_pop_million+factor(state)+factor(month),data=data_vmt_swaa)

rob.st2_st_main_swaa_trans<-coeftest(st2_st_main_swaa_trans,vcov = vcovHC(st2_st_main_swaa_trans,type = "HC"))

var_list <- c("swaa_wfh","st1_swaa_pred","st1_swaa_pred x transit","iv", "reopened", "gdp_p_in1000", "UR","vrm_19pct", "fare", "cases_p_1000", "dose_p_lag3","dose_p_lag3_6","dose_p_lag6_9","net_mig_rate", "l_pop_million")
var_list_new <- variable_mapping[var_list]


stargazer(ols_vmt_swaa,st2_st_main_swaa,st2_st_main_swaa_trans,st_state1_swaa,se = list(rob.ols_vmt_swaa[,"Std. Error"],rob.st2_st_main_swaa[,"Std. Error"],rob.st2_st_main_swaa_trans[,"Std. Error"],rob.st_state1_swaa[,"Std. Error"]),covariate.labels = var_list_new)

```


# 4. Plot time-varying effects
# 4.1 Summarize main effects by time
```{r}
library(zoo)
########### VMT
results_vmt <- tidy(rob.st2_st_time)
results_vmt2<-results_vmt%>%
           dplyr::rename(Variable = term,
                  Coefficient = estimate,
                  SE = std.error) %>% dplyr::select(-statistic,
                              -p.value)%>%
  mutate(conf.low_95=Coefficient-1.645*SE,conf.high_95=Coefficient+1.645*SE)

from <- as.Date('2020-04-01')
to <- as.Date('2022-09-01')

results_vmt3<-results_vmt2[2:11,]
results_vmt3$yq<-seq(as.yearqtr(from), as.yearqtr(to), by = 1/4)
results_vmt3$type<-'VMT'


########### transit
results_trans <- tidy(rob.st2_trans_time)
results_trans2<-results_trans%>%
           dplyr::rename(Variable = term,
                  Coefficient = estimate,
                  SE = std.error) %>% dplyr::select(-statistic,
                              -p.value)%>%
  mutate(conf.low_95=Coefficient-1.645*SE,conf.high_95=Coefficient+1.645*SE)

from <- as.Date('2020-04-01')
to <- as.Date('2022-09-01')

results_trans3<-results_trans2[2:11,]
results_trans3$yq<-seq(as.yearqtr(from), as.yearqtr(to), by = 1/4)
results_trans3$type<-'Transit ridership'

results_time_full<-rbind(results_trans3,results_vmt3)
results_time_full$type=factor(results_time_full$type,levels=c('Transit ridership','VMT'))
```

```{r}
results_time_vmt_trans<-results_time_full%>%
  filter(type%in%c('VMT', 'Transit ridership'))
results_time_vmt_trans
```
# 4.2 Plot time-varying effects
```{r}
ggplot(data=results_time_vmt_trans,aes(x = yq, 
                    y = Coefficient,group=type,color=type,shape=type))+ geom_smooth(method=lm,size=0.7,linetype='dashed',se=F)+
        geom_point(position=position_dodge(width = 0.1),size=3,stroke=1) + 
          geom_line(position=position_dodge(width = 0.1),lwd = 0.7) +
        geom_linerange(aes(x = yq,
                     ymin = conf.low_95,
                     ymax = conf.high_95),alpha=0.7,
                   lwd = 0.5,position=position_dodge(width = 0.1))+
scale_color_manual('',values=c("#F8766D","#00BA38"),labels=c('VMT', 'Transit ridership'),breaks=c('VMT', 'Transit ridership'))+
  scale_shape_manual('',values = c(0, 1),labels=c('VMT', 'Transit ridership'),breaks=c('VMT', 'Transit ridership'))+
   geom_hline(yintercept= 0,linetype='dashed')+
  zoo::scale_x_yearqtr(breaks = seq(from = min(results_time_full$yq), to = max(results_time_full$yq), by = 0.25),format = '%Y \nQ%q')+
  labs(x="")+
  theme(legend.position = c(0.6, 0.2))+
  theme(legend.key=element_blank(),
        # axis.ticks=element_blank(),
        axis.ticks.length=unit(.25, "cm"),
        axis.text=element_text(size=20,family="Arial"), 
        axis.text.x = element_text(vjust = 0.5, hjust=0.5),
        axis.title.y = element_text(size=25, family="Arial",margin = margin(r = 12)),
        axis.title.y.right = element_text(size=25, family="Arial",margin = margin(l = 12)),
        legend.box.background = element_rect(fill="transparent",colour=NA),
        legend.background = element_rect(fill='transparent'),
        legend.text=element_text(size=20, family="Arial",margin = margin(t = 3)), 
        panel.background=element_blank(),
        panel.border = element_rect(colour = 'black', fill=NA),
        panel.grid.major = element_line(colour = "lightgrey"), 
        panel.grid.minor = element_line(colour = "lightgrey"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        # legend.box.background = element_rect(colour = "black"),
        text = element_text(size = 20),
        plot.margin = margin(t = 0.5,  # Top margin
                             r = 1,  # Right margin
                             b = 0,  # Bottom margin
                             l = 0.5, "cm"))
# ggsave('plots/effect_time_vmt_trans.png', dpi = 800, width = 10, height = 7, units = "in")
```


